Hydras \& Co.: Formalized mathematics in Coq\\ for inspiration and entertainment

### 2021-11-18 doc/JFLA2022/paper.tex
pass over Gaia bridge section, mentioning some proof engineering concerns

\subsection{First steps towards a bridge to \gaia}

@@

In 2020, members of \community transferred the \gaia source code to GitHub and adapted it for recent releases of the Mathematical Components library, which \gaia heavily relies on. Anonymous volunteers (``collaborators of Nicolas Bourbaki'') then finished the only in-progress proof left by Grimm. At around 155,000 lines of code, \gaia is currently one of the largest maintained open source \coq projects~\cite{Gaia}.

\gaia contains definitions of ordinals in Cantor and Veblen normal form~\cite{grimm:hal-00911710}, adapted from the historical Cantor contribution~\cite{CantorContrib}. The data types for ordinals are essentially defined the same way in \gaia and \Hydras, but they are not identical inside \coq, e.g., due to residing in different modules. Moreover, there are small differences in the implementation of ordinal arithmetic due to different evolutionary paths taken since divergence from the ancestral library.
%
Nevertheless, we were able to establish an initial correspondence between both implementations of ordinals through rewriting lemmas. For instance, we proved that multiplication of ordinals in Cantor normal form in \Hydras refine \gaia's corresponding multiplication operation, and were able to import \gaia's proof of associativity of the multiplication almost for free.

@@

However, Dune is still inconvenient for local IDE-based Coq code development and does not yet support building Coqdoc files.
%
Therefore, we still support building the whole project using \make via the \coqmakefile tool bundled with \coq, and we rely on this build system in our documentation generation pipeline.
%
We expect to migrate fully to Dune as soon as these limitations are lifted.

>>>>

\subsection{Towards a bridge to \gaia}

@@

In 2020, members of \community transferred the \gaia source code to GitHub and adapted it for recent releases of the Mathematical Components library, which \gaia heavily relies on. Anonymous volunteers (``collaborators of Nicolas Bourbaki'') then finished the only in-progress proof left by Grimm. At around 155,000 LOC, \gaia is currently one of the largest maintained open source \coq projects~\cite{Gaia}.

\gaia contains definitions of ordinals in Cantor and Veblen normal form~\cite{grimm:hal-00911710}, adapted from the historical Cantor contribution~\cite{CantorContrib}. The data types for ordinals are essentially defined the same way as in \Hydras, but they are not identical inside \coq, e.g., due to residing in different modules. There are also minor differences in how ordinal arithmetic is implemented, due to the different evolutionary paths taken since divergence from the ancestor library.

As an initial step in bridging \gaia and \Hydras, we were able to establish a correspondence between both implementations of ordinals through rewriting lemmas. For instance, we proved that multiplication of ordinals in Cantor normal form in \Hydras refine \gaia's corresponding multiplication operation, and then imported \gaia's proof of associativity of the multiplication almost for free.

The initial bridge code uses the SSReflect proof language and idioms from the Mathematical Components library. We made this design decision since we believe it is less challenging to reason about Hydra-battles code using SSReflect and MathComp than to reason about Gaia without SSReflect.

@@

However, Dune is currently inconvenient for local IDE-based Coq code development and does not yet support building \coqdoc files.
%
Therefore, we still support building the whole project using \make via the \coqmakefile tool bundled with \coq, and we rely on this build system in our documentation generation pipeline.
%
We expect to migrate fully to Dune as soon as these limitations are lifted.
